parental migration and adolescents educational performance

### 2022-06-23 data/Gu-thesis-script.Rmd
First draft

Load packages
```{r}
library(tidyverse) 
library(haven)
library(fixest)
library(modelsummary)
library(MatchIt)
#library(vtable)
library(ggplot2)
library(gtsummary)
#library(DescTools)
```

@@

```{r}
d_dpvar <- d_migr %>%
  select(migr, 
         cog_dif, cog3pl, w2cog3pl, ttl_dif, w1ttl_100, w2ttl_100) %>%
  rename(
    `Cognitive score at wave 1` = cog3pl,
    `Cognitive score at wave 2` = w2cog3pl,
    `Academic exam score at wave 1` = w1ttl_100,
    `Academic exam score at wave 2` = w2ttl_100,
    `Cognitive score difference` = cog_dif,
    `Academic exam score difference` = ttl_dif
  ) 
```

@@

```{r}
d_dpvar <- as.data.frame(d_dpvar)
table_dependent_variable <- datasummary(
  All(d_dpvar) ~ Mean + SD + Min + Max,
  data = d_dpvar,
  output = 'markdown'
  )
```

```{r}
table_independent_variable <- datasummary_skim(
    data=d_idpvar,
    type = 'categorical',
    output = 'markdown')

@@

```{r}
set.seed(1)
fig_cog <- 
  ggplot(d_migr, aes(x='Parental Migration Type', y=cog_dif, fill = `Parental Migration`)) + 
    geom_jitter(color="black", size=0.4, alpha=0.9) +
    geom_boxplot(alpha=0.2)
```

```{r}
set.seed(1)
fig_ttl <- 
ggplot(d_migr, aes(x='Parental Migration Type', y=ttl_dif, fill = `Parental Migration`)) + 
    geom_jitter(color="black", size=0.4, alpha=0.9) +
    geom_boxplot(alpha=0.2)
```

```{r}
set.seed(1)
fig_cog_by_migration_type <-
  ggplot(d_migr, aes(x='Parental Migration Type', y=cog_dif, fill = `Parental Migration Type`)) + 
    geom_jitter(color="black", size=0.4, alpha=0.9) +
    geom_boxplot(alpha=0.2)

```{r}
set.seed(1)
fig_ttl_by_migration_type <-
  ggplot(d_migr, aes(x='Parental Migration Type', y=ttl_dif, fill = `Parental Migration Type`)) + 
    geom_jitter(color="black", size=0.4, alpha=0.9) +
    geom_boxplot(alpha=0.2)
```

@@

```{r}
table_cog <- models_cog %>% msummary(
  stars =  c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
  fmt = 2,
  title = 'Parental migration’s effect on children’s cognitive abilities',
  gof_omit = 'R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors',
  coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "cog3pl" = "W1 score"), 
  output = 'markdown')
table_cog
```

@@

```{r}
table_ttl <- models_ttl %>% msummary(
  stars =  c('†' = .1, '*' = .05, '**' = .01, '***' = .001), 
  title = 'Parental migration’s effect on children’s academic abilities',
  fmt = 2,
  coef_rename = c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_100" = "W1 score"), 
  gof_omit = 'R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors',
  output = 'markdown')
table_ttl
```

@@

```{r}
table_adj_cog <- models_adj_cog %>% msummary(
  stars =  c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           title = 'Parental migration’s effect on children’s cognitive abilities',
           fmt = 2, 
           coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "cog3pl" = "W1 score"),
  gof_omit = 'R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors',
           output = 'markdown'
  )
table_adj_cog
```

@@

```{r}
table_adj_ttl <- models_adj_ttl %>% msummary(
  stars =  c('†' = .1, '*' = .05, '**' = .01, '***' = .001), 
  title = 'Parental migration’s effect on children’s academic abilities',
  fmt = 2, 
  gof_omit = 'R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors',
  coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_100" = "W1 score"), output = 'markdown')
table_adj_ttl
```

@@

## Matching
```{r}
m.out <- matchit(
  migr ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_cl, 
  replace = T,
  ratio = 3,
  method = "nearest", 
  distance = "glm",
  exact = migr ~ ctyids
  )
summary(m.out) %>% plot()
```

@@

```{r}
d_mo_only <- d_cl %>% drop_na(migr_mo_only)
m.out_mo_only <- 
  matchit(migr_mo_only ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_mo_only, 
  replace = T,
  method = "nearest",
  distance = "glm",
  ratio = 3,
  exact = migr_mo_only ~ ctyids
  )
summary(m.out_mo_only) %>% plot()
```

@@

```{r}
d_fa_only <- d_cl %>% drop_na(migr_fa_only)
m.out_fa_only <- matchit(
  migr_fa_only ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet,
  data = d_fa_only,
  replace = T,
  method = "nearest",
  distance = "glm",
  ratio = 3,
  exact = migr_fa_only ~ ctyids
  )
summary(m.out_fa_only) %>% plot()
```

@@

```{r}
d_both <- d_cl %>% drop_na(migr_both) 
m.out_both <- matchit(
  migr_both ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level + home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_both, 
  method = "nearest",
  distance = "glm",
  replace = T,
  ratio = 3,
  exact = migr_both ~ ctyids
  )
summary(m.out_both) %>% plot()
```

@@

```{r}
models_m_cog <- list(
'Any Parent Absent' = didm_cog,
'Only Mother Absent' = didm_cog_mo_only,
##'Mother Absent' = #did_adj_cog_m,
'Only Father Absent' = didm_cog_fa_only,
##'Father Absent' = #did_adj_cog_f,
'Both Parents Absent' = didm_cog_both) 
table_m_cog <- models_m_cog %>% msummary(
  stars =  c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           title = 'Parental migration’s effect on children’s cognitive abilities',
           fmt = 2,
           gof_omit = 'R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors',
           coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "cog3pl" = "W1 score"), 
           output = 'markdown'
  )
table_m_cog
```

```{r}
models_m_ttl <- list(
'Any Parent Absent' = didm_ttl,
'Only Mother Absent' = didm_ttl_mo_only,
##'Mother Absent' = #did_adj_ttl_m,
'Only Father Absent' = didm_ttl_fa_only,
##'Father Absent' = #did_adj_ttl_f,
'Both Parents Absent' = didm_ttl_both) 
table_m_ttl <- models_m_ttl %>% msummary(
  stars =  c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           title = 'Parental migration’s effect on children’s academic abilities',
           fmt = 2,
           gof_omit = 'R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors',
           coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_100" = "W1 score"), 
           output = 'markdown'
  )
table_m_ttl
```

```{r}
models_m_adj_cog <- list(
'Any Parent Absent' = didm_adj_cog,
'Only Mother Absent' = didm_adj_cog_mo_only,
##'Mother Absent' = #did_adj_cog_m,
'Only Father Absent' = didm_adj_cog_fa_only,
##'Father Absent' = #did_adj_cog_f,
'Both Parents Absent' = didm_adj_cog_both) 
table_m_adj_cog <- models_m_adj_cog %>% msummary(
  stars =  c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           title = 'Parental migration’s effect on children’s cognitive abilities',
           fmt = 2,
           gof_omit = 'R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors',
           coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "cog3pl" = "W1 score"), 
           output = 'markdown'
  )
table_m_adj_cog
```

```{r}
models_m_adj_ttl <- list(
'Any Parent Absent' = didm_adj_ttl,
'Only Mother Absent' = didm_adj_ttl_mo_only,
##'Mother Absent' = #did_adj_ttl_m,
'Only Father Absent' = didm_adj_ttl_fa_only,
##'Father Absent' = #did_adj_ttl_f,
'Both Parents Absent' = didm_adj_ttl_both) 
table_m_adj_ttl <- models_m_adj_ttl %>% msummary(
  stars =  c('†' = .1, '*' = .05, '**' = .01, '***' = .001),
           title = 'Parental migration’s effect on children’s academic abilities',
           fmt = 2,
           gof_omit = 'R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors',
           coef_rename =  c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "w1ttl_100" = "W1 score"), 
           output = 'markdown'
  )
table_m_adj_ttl
```

@@

```{r}
d_cl1=d_cl1 %>% mutate(
  supervision.less = if_else(b2201 > w2a18 & b2202 > w2a19, 1, 0)
)
```

>>>>

Load packages
```{r}
library(tidyverse) 
library(haven)
library(fixest)
library(modelsummary)
library(MatchIt)
#library(vtable)
library(ggplot2)
#library(gtsummary)
#library(DescTools)
#knitr::opts_chunk$set(modelsummary_get="broom")
```

@@

```{r}
d_dpvar <- d_migr %>%
  select(#migr, 
         cog_dif, cog3pl, w2cog3pl, ttl_dif, w1ttl_100, w2ttl_100) %>%
  rename(
    `Cognitive score at wave 1` = cog3pl,
    `Cognitive score at wave 2` = w2cog3pl,
    `Academic exam score at wave 1` = w1ttl_100,
    `Academic exam score at wave 2` = w2ttl_100,
    `Cognitive score difference` = cog_dif,
    `Academic exam score difference` = ttl_dif
  ) 
```

@@

```{r}
d_dpvar <- as.data.frame(d_dpvar)
table_dependent_variable <- datasummary(
  All(d_dpvar) ~ Mean + SD + Min + Max,
  data = d_dpvar,
  output = 'markdown'
  )
table_dependent_variable
```

```{r}
table_independent_variable <- datasummary_skim(
    data=d_idpvar,
    type = 'categorical',
    output = 'markdown')
table_independent_variable
```

@@

```{r}
set.seed(1)
fig_cog <- 
  ggplot(d_migr, aes(x=`Parental Migration`, y=cog_dif, fill = `Parental Migration`)) + 
    geom_violin()+
    xlab('Parental Migration') +
    theme(legend.position="none") +
    ylab("")
fig_cog
```

```{r}
set.seed(1)
fig_ttl <- 
  ggplot(d_migr, aes(x=`Parental Migration`, y=ttl_dif, fill = `Parental Migration`)) + 
    geom_violin()+
    xlab('Parental Migration') +
    theme(legend.position="none") +
    ylab("")
fig_ttl
```

```{r}
set.seed(1)
fig_cog_by_migration_type <-
  ggplot(d_migr, aes(x=`Parental Migration Type`, y=cog_dif, fill = `Parental Migration Type`)) + 
    geom_violin()+
    xlab('Parental Migration Type') +
    theme(legend.position="none") +
    ylab("")
fig_cog_by_migration_type
```

```{r}
set.seed(1)
fig_ttl_by_migration_type <-
  ggplot(d_migr, aes(x=`Parental Migration Type`, y=ttl_dif, fill = `Parental Migration Type`)) + 
    geom_violin()+
    xlab('Parental Migration Type') +
    theme(legend.position="none") +
    ylab("")
fig_ttl_by_migration_type
```

@@

```{r}
stars <- c('†' = .1, '*' = .05, '**' = .01, '***' = .001)
coef_rename <- c("migr" = "Explanator", "migr_mo_only" = "Explanator", "migr_mo" = "Explanator", "migr_fa" = "Explanator", "migr_fa_only" = "Explanator", "migr_both" = "Explanator", "cog3pl" = "W1 score", "w1ttl_100" = "W1 score")
title_cog <- 'Parental migration’s effect on children’s cognitive abilities'
title_ttl <- 'Parental migration’s effect on children’s academic abilities'
gof_omit <- 'R2 Pseudo|AIC|BIC|Log.Lik.|Std.Errors'
```

```{r}
table_cog <- models_cog %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'markdown')
table_cog
```

@@

```{r}
table_ttl <- models_ttl %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'markdown')
table_ttl
```

@@

```{r}
table_adj_cog <- models_adj_cog %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'markdown'
  )
table_adj_cog
```

@@

```{r}
table_adj_ttl <- models_adj_ttl %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  output = 'markdown'
  )
table_adj_ttl
```

@@

## Matching
```{r}
m.out <- matchit(
  migr ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_cl, 
  replace = T,
  ratio = 3,
  method = "nearest", 
  distance = "glm",
  exact = migr ~ ctyids
  )
fig_msmd_any <- summary(m.out) %>% plot(xlim = c(0, 0.55), position = 'right')
```

@@

```{r}
d_mo_only <- d_cl %>% drop_na(migr_mo_only)
m.out_mo_only <- 
  matchit(migr_mo_only ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_mo_only, 
  replace = T,
  method = "nearest",
  distance = "glm",
  ratio = 3,
  exact = migr_mo_only ~ ctyids
  )
fig_msmd_mo_only <- summary(m.out_mo_only) %>% plot(xlim = c(0, 0.55), position = 'right')
```

@@

```{r}
d_fa_only <- d_cl %>% drop_na(migr_fa_only)
m.out_fa_only <- matchit(
  migr_fa_only ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level +home.economic.resource + home.extracurricular.book + home.has.internet,
  data = d_fa_only,
  replace = T,
  method = "nearest",
  distance = "glm",
  ratio = 3,
  exact = migr_fa_only ~ ctyids
  )
fig_msmd_fa_only <- summary(m.out_fa_only) %>% plot(xlim = c(0, 0.55), position = 'right')
```

@@

```{r}
d_both <- d_cl %>% drop_na(migr_both) 
m.out_both <- matchit(
  migr_both ~ cog3pl + w1ttl_100 + child.age + child.is.girl + child.health.self.rated + child.has.rural.hukou + child.number.of.sibling + child.went.to.preschool + child.had.skipped.grade + child.had.repeated.grade + child.in.boarding.school + child.is.ethnic.minority + child.educational.aspiration + parent.expectation.on.child + parent.has.white.collar.job + parent.educational.level + home.economic.resource + home.extracurricular.book + home.has.internet, 
  data = d_both, 
  method = "nearest",
  distance = "glm",
  replace = T,
  ratio = 3,
  exact = migr_both ~ ctyids
  )
fig_msmd_both <- plot(summary(m.out_both), xlim = c(0, 0.55), position = 'right')
```

@@

```{r}
title_cog_matching <- 'Parental migration’s effect on children’s cognitive abilities, estimated with matching'
title_ttl_matching <- 'Parental migration’s effect on children’s academic abilities, estimated with matching'
note_matching <- "3:1 nearest neighbor propensity score matching with replacement; exact matching by county"
```

```{r}
models_m_cog <- list(
'Any Parent Absent' = didm_cog,
'Only Mother Absent' = didm_cog_mo_only,
##'Mother Absent' = #did_adj_cog_m,
'Only Father Absent' = didm_cog_fa_only,
##'Father Absent' = #did_adj_cog_f,
'Both Parents Absent' = didm_cog_both) 
```

```{r}
table_m_cog <- models_m_cog %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_cog_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'markdown'
  )
table_m_cog
```

```{r}
models_m_ttl <- list(
'Any Parent Absent' = didm_ttl,
'Only Mother Absent' = didm_ttl_mo_only,
##'Mother Absent' = #did_adj_ttl_m,
'Only Father Absent' = didm_ttl_fa_only,
##'Father Absent' = #did_adj_ttl_f,
'Both Parents Absent' = didm_ttl_both) 
table_m_ttl <- models_m_ttl %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'markdown'
  )
table_m_ttl
```

```{r}
models_m_adj_cog <- list(
'Any Parent Absent' = didm_adj_cog,
'Only Mother Absent' = didm_adj_cog_mo_only,
##'Mother Absent' = #did_adj_cog_m,
'Only Father Absent' = didm_adj_cog_fa_only,
##'Father Absent' = #did_adj_cog_f,
'Both Parents Absent' = didm_adj_cog_both) 
table_m_adj_cog <- models_m_adj_cog %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'markdown'
  )
table_m_adj_cog
```

```{r}
models_m_adj_ttl <- list(
'Any Parent Absent' = didm_adj_ttl,
'Only Mother Absent' = didm_adj_ttl_mo_only,
##'Mother Absent' = #did_adj_ttl_m,
'Only Father Absent' = didm_adj_ttl_fa_only,
##'Father Absent' = #did_adj_ttl_f,
'Both Parents Absent' = didm_adj_ttl_both) 
table_m_adj_ttl <- models_m_adj_ttl %>% msummary(
  stars =  stars,
  fmt = 2,
  title = title_ttl_matching,
  gof_omit = gof_omit,
  coef_rename =  coef_rename, 
  notes = note_matching,
  output = 'markdown'
  )
table_m_adj_ttl
```

@@

```{r}
d_cl1=d_cl1 %>% mutate(
  supervision.less = if_else((b2201 + b2202) > (w2a18 + w2a19), 1, 0)
)
```
